A coherent and comprehensive dating system for stars in the Milky Way

### 2017-05-08 paper/ms.tex
adding words and equations to methods section

\documentclass[useAMS, usenatbib, preprint, 12pt]{aastex}
\usepackage{cite, natbib}
\usepackage{float}
\usepackage{epsfig}
\usepackage{cases}
\usepackage[section]{placeins}
\usepackage{graphicx, subfigure}
\usepackage{color}
\usepackage{bm}
\usepackage{enumerate}
\usepackage{pdflscape}

@@

% Stellar parameters
\newcommand{\Teff}{$T_{\mathrm{eff}}$}
\newcommand{\teff}{$T_{\mathrm{eff}}$}
\newcommand{\FeH}{[Fe/H]}
\newcommand{\feh}{[Fe/H]}
\newcommand{\logg}{log \emph{g}}
\newcommand{\dnu}{$\Delta \nu$}
\newcommand{\numax}{$\nu_{\mathrm{max}}$}

@@

\begin{abstract}
% Aims
Stellar age underpins a wide range of astronomical applications but
remains one of the most challenging stellar properties to infer.
% Method
Competing dating methods with different observational inputs often produce
    wildly inconsistent age estimates.
In this work we attempt to combine multiple methods into one coherent model.
For the first time, we combine apparent magnitudes and colour information,
    typically used to infer an age using stellar evolutionary tracks, with
    other age proxies: rotation period and galactic kinematics.
We identify a sample of stars with rotation periods, colours, spectroscopic
    parameters and kinematics to calibrate this dating method.
% Results
\end{abstract}

@@

We combine the following dating methods: rotation, stellar evolution models
and kinematics into one model via self calibration.
As an example, take the Sun: a star with temperature $5777K$, $ \log(g) =
4.44$ $[Fe/H] = 0.0$ and $P_{rot} = 26$ days.

@@

% acknowledgements
This research was funded by the Simons Foundation.
Some of the data presented in this paper were obtained from the Mikulski
Archive for Space Telescopes (MAST).
STScI is operated by the Association of Universities for Research in
Astronomy, Inc., under NASA contract NAS5-26555.
Support for MAST for non-HST data is provided by the NASA Office of Space
Science via grant NNX09AF08G and by other grants and contracts.
This paper includes data collected by the Kepler mission. Funding for the
Kepler mission is provided by the NASA Science Mission directorate.
\racomment{Add TGAS \& Apogee acknowledgements.}

>>>>

\documentclass[useAMS, usenatbib, preprint, 12pt]{aastex}
\usepackage{cite, natbib}
\usepackage{float}
\usepackage{epsfig}
\usepackage{cases}
\usepackage[section]{placeins}
\usepackage{graphicx, subfigure}
\usepackage{color}
\usepackage{bm}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{eqnarray}

@@

% Stellar parameters
\newcommand{\Teff}{$T_{\mathrm{eff}}$}
\newcommand{\teff}{$T_{\mathrm{eff}}$}
\newcommand{\FeH}{[Fe/H]}
\newcommand{\feh}{[Fe/H]}
\newcommand{\Av}{$A_v$}
\newcommand{\av}{$A_v$}
\newcommand{\logg}{log \emph{g}}
\newcommand{\dnu}{$\Delta \nu$}
\newcommand{\numax}{$\nu_{\mathrm{max}}$}

@@

\begin{abstract}
% Aims
Stellar age underpins a wide range of astronomical applications but
remains one of the most challenging stellar properties to infer.
% Method
Competing dating methods with different observational inputs often produce
    wildly inconsistent age estimates.
In this work we attempt to combine multiple methods into one coherent model.
For the first time, we combine apparent magnitudes and colour information,
    typically used to infer an age using stellar evolutionary tracks, with
    other age proxies: rotation period and vertical action.
We identify a sample of stars with rotation periods, colours, spectroscopic
    parameters and kinematics to calibrate this dating method.
% Results
\end{abstract}

@@

We combine the following dating methods: gyrochronology, isochrone fitting,
asteroseismology and galactic kinematics into one model.
In classical isochrone fitting, one might calculate the probability of age
($A$), mass ($M$), \feh, distance ($D$) and extinction (\av), given a set of
apparent magnitudes in different filters, \eg\ $M_V$, $M_J$ and $M_K$
\begin{equation}
    p(A, M, [Fe/H], D, A_v | M_V, M_J, M_K) \propto p(M_v, M_J, M_K |
    A, M, [Fe/H], D, A_v)p(A, M, [Fe/H], D, A_v).
\end{equation}
Note that there are no free parameters in the model here --- the model is
entirely governed by physics and does not need to be calibrated further.
Similarly, in classical gyrochronology, one might calculate the probability of
age given rotation period and $B-V$ colour,
\begin{equation}
    p(A | P_{rot}, B-V) \propto p(P_{rot}, B-V | A)p(A)
\end{equation}
In a gyrochronology model, because our understanding of the physics is
incomplete, there are some free parameters that must be calibrated.
In the \citet{Barnes2003} gyrochronology parameterisation, these are $a$, $b$
and $n$, $A = P^{1/n} - (B-V-0.4)^{b/n}$.
During a calibration exercise, the observables, $A$, $P_{rot}$ and $B-V$ will
be known and the parameters, $a$, $b$ and $n$ will be inferred,
\begin{equation}
    p(a, b, n | A, P_{rot}, B-V) \propto p(A, P_{rot}, B-V | a,
    b, n) p(a, b, n).
\end{equation}
The vertical action-age relation also has free parameters that must be
calibrated using observations, $A = \alpha (\sigma_{Jz}^2)^\beta$.
These free parameters, $\alpha$ and $\beta$ can, again be calibrated via
\begin{equation}
    p(\alpha, \beta | \{A\}, \{J_z\}) \propto p(\{A\}, \{J_z\} | \alpha,
    \beta) p(\alpha, \beta).
\end{equation}

The unique aspect of our analysis is to combine inference over ages with
calibration of the free parameters in the gyrochronology and vertical
action-age relations:
\begin{eqnarray}
    &&p(\{A\}, \{M\}, \{[Fe/H]\}, \{D\}, \{A_v\}, a, b, n, \alpha, \beta|
    \{M_V\}, \{M_J\}, \{M_K\}, \{P_{rot}\}, \{B-V\}, \{J_z\}) \\
    \propto &&p(\{M_v\}, \{M_J\}, \{M_K\}, \{P_{rot}\},
    \{B-V\}, \{J_z\} | \{A\}, \{M\}, \{[Fe/H]\}, \{D\}, \{A_v\}, a, b, n,
    \alpha, \beta) \\
    &&p(A, M, [Fe/H], D, A_v, a, b, n, \alpha, \beta).
\end{eqnarray}
We can use conditional independence to break this down further into the
following:
\begin{eqnarray}
    &&p(\{A\}, \{M\}, \{[Fe/H]\}, \{D\}, \{A_v\}, a, b, n, \alpha, \beta|
    M_V, M_J, M_K,
    \{P_{rot}\}, \{B-V\}, \{J_z\}) \\
    \propto &&p(\{M_v\}, \{M_J\}, \{M_K\} | \{A\}, \{M\}, \{[Fe/H]\}, \{D\},
    \{A_v\})
    p(A)p(M)p([Fe/H])p(D)p(A_v) \\
    &&p(\{P_{rot}\}, \{B-V\}, a, b, n | A)p(a)p(b)p(n) \\
    &&p(\{J_z\}, \alpha, \beta | \{A\}) p(\alpha)p(\beta).
\end{eqnarray}
In other words
\begin{eqnarray}
    &&p(\mathrm{Ages~\&~other~parameters}|\mathrm{observables}) \\
    \propto &&\mathcal{L}_{\mathrm{isochronal~age}} \times
    \mathcal{L}_{\mathrm{gyro~age}} \times
    \mathcal{L}_{\mathrm{kinematic~age}} \\
    && \times p(\mathrm{Ages~\&~other~parameters})
\end{eqnarray}.

As an example, take the Sun: a star with temperature $5777K$, $ \log(g) =
4.44$ $[Fe/H] = 0.0$ and $P_{rot} = 26$ days.

@@

% acknowledgements
This research was funded by the Simons Foundation.
Some of the data presented in this paper were obtained from the Mikulski
Archive for Space Telescopes (MAST).
STScI is operated by the Association of Universities for Research in
Astronomy, Inc., under NASA contract NAS5-26555.
Support for MAST for non-HST data is provided by the NASA Office of Space
Science via grant NNX09AF08G and by other grants and contracts.
This paper includes data collected by the Kepler mission. Funding for the
Kepler mission is provided by the NASA Science Mission directorate.
\racomment{Add Gaia sprint, TGAS \& Apogee acknowledgements.}
